---
title: "Egg Analysis"
author: "Ana Dulskiy"
date: "4/1/2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries:

```{r}
library(tidyverse)
library(patchwork)
```

Load data:

getwd()
setwd("~/egg-biochem")

```{r}
egg <- read.csv("eggdata_clean.csv")
```

## Plots!

### Egg Diameter
```{r}
diameter <- ggplot(egg, aes(x = species, y = diameter, color = population)) +
  geom_boxplot()

```



### Egg volume

```{r}
volume <- ggplot(egg, aes(x = species, y = volume, color = population)) +
  geom_boxplot()
```


### Total Protein

```{r}
total.protein <- ggplot(egg, aes(x = species, y = total.protein, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

### Protein Density

```{r}
protein.density <- ggplot(egg, aes(x = species, y = protein.density, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")

```


### Total Lipid

```{r}
total.lipid <- ggplot(egg, aes(x = species, y = total.lipid, color = population)) +
  geom_boxplot()  +
  theme(legend.position = "none")
```

### Lipid Density

```{r}
lipid.density <- ggplot(egg, aes(x = species, y = lipid.density, color = population)) +
  geom_boxplot()  +
  theme(legend.position = "none")
```

### Total Carb

```{r}
total.carb <- ggplot(egg, aes(x = species, y = total.carb, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Carb Density
```{r}
carb.density <- ggplot(egg, aes(x = species, y = carb.density, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Protein Energy
```{r}
protein.energy <- ggplot(egg, aes(x = species, y = protein.energy, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Lipid Energy
```{r}
lipid.energy <- ggplot(egg, aes(x = species, y = lipid.energy, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Carb Energy
```{r}
carb.energy <- ggplot(egg, aes(x = species, y = carb.energy, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Total Energy
```{r}
total.energy <- ggplot(egg, aes(x = species, y = total.energy, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Energy Density
```{r}
energy.density <- ggplot(egg, aes(x = species, y = energy.density, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```


### Lipid:Protein
```{r}
lipid.protein <- ggplot(egg, aes(x = species, y = lipid.protein, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

### % Energy from Lipid
```{r}
perc.energy.lipid <- ggplot(egg, aes(x = species, y = perc.energy.lipid, color = population)) +
  geom_boxplot()  +
  theme(legend.position = "none")
```
### % Protein
```{r}
perc.protein <- ggplot(egg, aes(x = species, y = perc.protein, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

### % Lipid
```{r}
perc.lipid <- ggplot(egg, aes(x = species, y = perc.lipid, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

### % Carb
```{r}
perc.carb <- ggplot(egg, aes(x = species, y = perc.carb, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

### Summed Constituents

```{r}
summed.constituents <- ggplot(egg, aes(x = species, y = summed.constituents, color = population)) +
  geom_boxplot() +
  theme(legend.position = "none")
```

Compiled plots:
```{r}
diameter + volume
```
```{r}
total.protein + total.lipid + total.carb
```
```{r}
protein.density + lipid.density + carb.density
```
```{r}
protein.energy + lipid.energy + carb.energy
```

```{r}
perc.protein + perc.lipid + perc.carb
```
```{r}
perc.energy.lipid + lipid.protein
```
```{r}
total.energy + summed.constituents
```


# Statistics

### Summary:
```{r}
summary(egg)
```

### 2-way ANOVA unbalanced
```{r}
install.packages("car")
library(car)
install.packages("fitdistrplus")
library(fitdistrplus)
```
Check if data normal (p-value >0.05 for shapiro test = normal):
```{r}
lshap <- lapply(egg[5:23], shapiro.test)
lres <- sapply(lshap, `[`, c("statistic","p.value"))
t(lres)

```

The following variables are not normally distributed according to the shapiro test and have to be treated accordingly:
- Diameter
- Volume
- Total Carb
- Carb Energy


Figure out what distributions we're dealing with for non-normal data:
```{r}
library(fitdistrplus)
descdist(egg$diameter, discrete = FALSE)

```
Diameter = beta distribution

```{r}
descdist(egg$volume, discrete = FALSE)

```

Volume = beta distribution

```{r}
#total.carb has one NA value, so we need to remove that to do the descdist plot
total.carb.nn <- c(na.exclude(egg$total.carb))
descdist(total.carb.nn, discrete = FALSE)

```
Total carb = beta distribution

```{r}
carb.energy.nn <- c(na.exclude(egg$carb.energy))
descdist(carb.energy.nn, discrete = FALSE)
```
Carb energy = beta distribution

All non-normal variables = beta distributions!


### Beta regression models (diameter, volume, total carb, carb energy)

```{r}
install.packages("betareg")
library(betareg)
volume.beta <- betareg(volume ~ population*species, data = egg)
summary(volume.beta)
egg_stats <- 
  egg_stats %>%
  mutate(volume_pred = predict(volume.beta))

```










```{r}
volume.mod <- lm(volume ~ population * species, egg)

plot(volume.mod, which = 2, add.smooth = FALSE)
plot(volume.mod, which = 3, add.smooth = FALSE)

```
```{r}
library(MuMIn)
total.protein.mod <- glm(total.protein ~ population * species, data = egg)
summary(total.protein.mod)

plot(total.protein.mod, which = 2, add.smooth = FALSE)
plot(total.protein.mod, which = 3, add.smooth = FALSE)

```


```{r}
anova(lm(volume ~ species * population, egg))

```


### Modeling

library(MuMIn)

Check if data normal (p-value >0.05 for shapiro test = normal):
```{r}
lshap <- lapply(egg[5:23], shapiro.test)
lres <- sapply(lshap, `[`, c("statistic","p.value"))
t(lres)
```

The following variables are not normally distributed and have to be treated accordingly:
- Diameter
- Volume
- Total Carb
- Carb Energy

```{r}
vol.dist <- ggplot(egg, aes(x = X, y = volume)) +
  geom_line()

diam.dist <- ggplot(egg, aes(x = X, y = diameter)) +
  geom_line()

totcarb.dist <- ggplot(egg, aes(x = X, y = total.carb)) +
  geom_line()

carbenergy.dist <- ggplot(egg, aes(x = X, y = carb.energy)) +
  geom_line()


vol.dist + diam.dist + totcarb.dist + carbenergy.dist
```
```{r}

```


GLMs
```{r}
total.protein.glm <- glm(total.protein ~ population * species, data = egg)
anova(total.protein.glm, test = "Chisq")
summary(total.protein.glm)
egg_stats <- egg %>%
  mutate(total.protein.glm = predict(total.protein.glm))


```

```{r}

```

