---
title: "egg-analysis-2"
author: "Ana Dulskiy"
date: "4/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Load libraries:

```{r}
library(tidyverse)
library(patchwork)
library(MuMIn)
library(car)
library(fitdistrplus)
library(betareg)
```

Load data:
```{r}
getwd()
egg <- read.csv("~/eggdata_clean.csv")
```

### 2-way ANOVA unbalanced

Check if data normal (p-value >0.05 for shapiro test = normal):
```{r}
lshap <- lapply(egg[4:22], shapiro.test)
lres <- sapply(lshap, `[`, c("statistic","p.value"))
t(lres)

```

The following variables are not normally distributed according to the shapiro test and have to be treated accordingly:
- Diameter
- Volume
- Total Carb
- Carb Energy


Figure out what distributions we're dealing with for non-normal data:
```{r}
descdist(egg$diameter, discrete = FALSE, boot = 100)
```
Diameter = beta distribution

```{r}
descdist(egg$volume, discrete = FALSE)

```

Volume = beta distribution

```{r}
#total.carb has one NA value, so we need to remove that to do the descdist plot
total.carb.nn <- c(na.exclude(egg$total.carb))
descdist(total.carb.nn, discrete = FALSE)

```
Total carb = beta distribution

```{r}
carb.energy.nn <- c(na.exclude(egg$carb.energy))
descdist(carb.energy.nn, discrete = FALSE)
```
Carb energy = beta distribution

All non-normal variables = beta distributions!


### Beta regression models (diameter, volume, total carb, carb energy)

```{r}
diameter.beta <- betareg(diameter ~ population*species, data = egg)
summary(diameter.beta)
# doesn't work on diameter... maybe beta isn't the right call
```

```{r}
volume.beta <- betareg(volume ~ population*species, data = egg)
summary(volume.beta)
egg_stats <- 
  egg %>%
  mutate(volumepred = predict(volume.beta))

```


### GLMs (Gaussian)

Total Protein:
```{r}
total.protein.mod <- glm(total.protein ~ population*species, data = egg)
summary(total.protein.mod)
egg_stats <- egg_stats %>%
  mutate(total.protein.pred = predict(total.protein.mod))
```



Protein Density:
```{r}
protein.density.mod <- glm(protein.density ~ population*species, data = egg)
summary(protein.density.mod)
egg_stats <- egg_stats %>%
  mutate(protein.density.pred = predict(protein.density.mod))
```

Total Lipid:
```{r}
total.lipid.mod <- glm(total.lipid ~ population*species, data = egg)
summary(total.lipid.mod)
egg_stats <- egg_stats %>%
  mutate(total.lipid.pred = predict(total.lipid.mod))
```


Lipid Density:
```{r}
lipid.density.mod <- glm(lipid.density ~ population*species, data = egg)
summary(lipid.density.mod)
egg_stats <- egg_stats %>%
  mutate(lipid.density.pred = predict(lipid.density.mod))
```

Total Carb:
```{r}
total.carb.mod <- glm(total.carb ~ population*species, data = egg)
summary(total.carb.mod)
egg_stats <- egg_stats %>%
  mutate(total.carb.pred = predict(total.carb.mod))
```

Carb Density:
Protein Energy:
Lipid Energy:
Carb Energy:
Total:Energy





```{r}
volume.mod <- lm(volume ~ population * species, egg)

plot(volume.mod, which = 2, add.smooth = FALSE)
plot(volume.mod, which = 3, add.smooth = FALSE)

```


```{r}

total.protein.mod <- glm(total.protein ~ population * species, data = egg)
summary(total.protein.mod)

plot(total.protein.mod, which = 2, add.smooth = FALSE)
plot(total.protein.mod, which = 3, add.smooth = FALSE)

```








### Modeling

library(MuMIn)

```{r}
vol.dist <- ggplot(egg, aes(x = X, y = volume)) +
  geom_line()

diam.dist <- ggplot(egg, aes(x = X, y = diameter)) +
  geom_line()

totcarb.dist <- ggplot(egg, aes(x = X, y = total.carb)) +
  geom_line()

carbenergy.dist <- ggplot(egg, aes(x = X, y = carb.energy)) +
  geom_line()


vol.dist + diam.dist + totcarb.dist + carbenergy.dist
```
```{r}

```


GLMs
```{r}
total.protein.glm <- glm(total.protein ~ population * species, data = egg)
anova(total.protein.glm, test = "Chisq")
summary(total.protein.glm)
egg_stats <- egg %>%
  mutate(total.protein.glm = predict(total.protein.glm))


```

```{r}

```

